---
title: "Checksums"
author: "Ivan Calandra"
date: "`r Sys.time()`"
output:
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
---


```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---


# Goal of the script
This script calculates the MD5 checksums of all files in the repository to check for integrity and file versions.


---


# Load packages
```{r}
library(tools)
library(tidyverse)
```


---


# Get names, path and MD5 checksums of all files
## R analysis
```{r}
check_R <- list.files(path = "R_analysis", full.names = TRUE, recursive = TRUE) %>% 
           md5sum()
```

The MD5 checksums of the files created during the R analysis are:  
```{r, echo = FALSE}
data.frame(files = names(check_R), checksum = check_R, row.names = NULL)
```

## Python analysis
```{r}
check_Python <- list.files(path = "Python_analysis", full.names = TRUE, recursive = TRUE) %>% 
                md5sum()
```

The MD5 checksums of the files created during the Python analysis are:  
```{r, echo = FALSE}
data.frame(files = names(check_Python), checksum = check_Python, row.names = NULL)
```


---


# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("R_analysis/scripts/SSFA_0_RStudioVersion.txt", n = 1)`.


---


END OF SCRIPT
