---
title: "Plots of SSFA variables"
author: "Ivan Calandra"
date: "`r Sys.time()`"
output:
  github_document: 
    toc: true
    toc_depth: 2
    html_preview: false
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---


# Goal of the script
The script plots all SSFA variables for each dataset to compare between the ConfoMap and Toothfrax analyses.   

```{r}
dir_in  <- "R_analysis/derived_data"
dir_out <- "R_analysis/plots"
```

Input Rbin data file must be located in "`r paste0("~/", dir_in)`".  
Plots will be saved in "`r paste0("~/", dir_out)`".

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(ggh4x)
```


---


# Read in data
## Get name and path of input file 
```{r}
info_in <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
info_in
```


## Read in Rbin file
```{r}
all_data <- loadObject(info_in)
str(all_data)
```


---


# Define variables
```{r}
# x-axis (grouping)
x_var_GP <- x_var_sheep <- "Diet"
x_var_lith <- "Before.after"

# y-axis
y_var <- c("Asfc", "Smfc", "HAsfc9", "HAsfc81", "epLsar", "NewEplsar")

# colors
grp_colors <- "Software"

# subplots Lithics dataset
facet_lith <- "Treatment"
```

The following variables will be used:  
```{r}
x_var_GP
x_var_sheep
x_var_lith
grp_colors
facet_lith
y_var
```


---


# Calculate y-scales
The range of the y-scales on the plots should be the same for the guinea pig and sheep datasets. Lithics are not comparable at all so this dataset is plotted with appropriate y-scales for this dataset alone.

```{r}
# Select guinea pig and sheep datasets
GP_sheep <- filter(all_data, Dataset %in% c("GuineaPigs", "Sheeps"))

# Create a named empty list to store the ranges of each parameter
yscales <- vector(mode = "list", length = length(y_var))
names(yscales) <- y_var

# Calculate the range of each parameter
for (i in y_var) {
  yscales[[i]] <- scale_y_continuous(limits = range(GP_sheep[[i]], na.rm = TRUE))
}
```

---


# Number of diets
In order to have the same width of the boxes for the sheep and guinea pig datasets, we need to calculate how many diets there are in each dataset.
```{r}
diet_GP <- length(unique(all_data[all_data$Dataset == "GuineaPigs", "Diet"]))
diet_sheep <- length(unique(all_data[all_data$Dataset == "Sheeps", "Diet"]))
ratio_diet <- diet_GP / diet_sheep
```


---


# Plot each set of the selected numeric variables
## Guinea Pigs
```{r}
# Filter dataset
GP <- filter(all_data, Dataset == "GuineaPigs")

# Change from wide to long format
GP_plot <- pivot_longer(GP[c(x_var_GP, grp_colors, y_var)], all_of(y_var))

# Re-order factor levels to fit order of plots on facet
GP_plot$name <- factor(GP_plot$name, levels = y_var)

# Plot all variables at once using facet_wrap()
p_GP <- ggplot(GP_plot, aes_string(x = x_var_GP, y = "value", color = grp_colors)) +
        
        # Hide outliers (all points are shown with geom_point() below)
        # The factor 'ratio_diet' ensures that the boxes have the same width 
        # as for the sheep dataset
        geom_boxplot(outlier.shape = NA, 
                     position = position_dodge(preserve = "single", width = 0.9*ratio_diet),
                     width = 0.8*ratio_diet) + 
  
        # Define jitter for points within boxplots
        # The factor 'ratio_diet' ensures that the boxes have the same width 
        # as for the sheep dataset
        geom_point(position = position_jitterdodge(jitter.width = 0.3*ratio_diet,
                                                   dodge.width = 0.9*ratio_diet,
                                                   seed = 123)) +
  
        # Wrap around parameters
        facet_wrap(~name, scales = "free_y", ncol = 2) +
  
        # Use custom y-scales
        facetted_pos_scales(y = yscales) +
  
        # Remove axis labels
        labs(x = NULL, y = NULL) + 

        # Choose a light theme
        theme_classic() +

        # Move legend to bottom and remove legend title
        theme(legend.position="bottom", legend.title = element_blank())

# Print and save resulting plot
print(p_GP)
ggsave(filename = "/SSFA_GuineaPigs_plot.pdf", path = dir_out, 
       device = "pdf", width = 190, height = 240, units = "mm")
```

Warnings are due to the missing values for Toothfrax - NewEplsar.


## Sheeps
Note that for compactness, comments to explain the code are given only in the section about [Guinea Pigs](#guinea-pigs).

```{r}
sheep <- filter(all_data, Dataset == "Sheeps")
sheep_plot <- pivot_longer(sheep[c(x_var_sheep, grp_colors, y_var)], all_of(y_var))
sheep_plot$name <- factor(sheep_plot$name, levels = y_var)
p_sheep <- ggplot(sheep_plot, aes_string(x = x_var_sheep, y = "value", color = grp_colors)) +
           geom_boxplot(outlier.shape = NA, 
                        position = position_dodge(preserve = "single", width = 0.9),
                        width = 0.8) + 
           geom_point(position = position_jitterdodge(jitter.width = 0.3, 
                                                      dodge.width = 0.9,
                                                      seed = 123)) +
           facet_wrap(~name, scales = "free_y", ncol = 2) +
           facetted_pos_scales(y = yscales) +
           labs(x = NULL, y = NULL) + 
           theme_classic() +
           theme(legend.position="bottom", legend.title = element_blank())
print(p_sheep)
ggsave(filename = "/SSFA_Sheeps_plot.pdf", path = dir_out, 
       device = "pdf", width = 190, height = 240, units = "mm")
```


## Lithics
Note that for compactness, comments to explain the code are given only in the section about [Guinea Pigs](#guinea-pigs). 

There is one difference though: here, three columns are used for the grouping ("Software", "Treatment", and "Before.after").  
Software is still shown with colors.  
"Before.after" is plotted on the x-axis and the variables are on the y-axes.  
`facet_grid()` is used to plot a grid of subplots, with variables in rows and treatments in columns.

```{r}
lith <- filter(all_data, Dataset == "Lithics")
lith_plot <- pivot_longer(lith[c(x_var_lith, grp_colors, facet_lith, y_var)], all_of(y_var))
lith_plot$name <- factor(lith_plot$name, levels = y_var)
p_lith <- ggplot(lith_plot, aes_string(x = x_var_lith, y = "value", color = grp_colors)) +
          geom_boxplot(outlier.shape = NA,
                       position = position_dodge(preserve = "single", width = 0.8),
                       width = 0.7) + 
          geom_point(position = position_jitterdodge(jitter.width = 0.3, 
                                                     dodge.width = 0.8,
                                                     seed = 123)) +
          facet_grid(as.formula(paste0("name~", facet_lith)), scales = "free_y") +
          labs(x = NULL, y = NULL) + 
          theme_classic() +
          theme(legend.position="bottom", legend.title = element_blank())
print(p_lith)
ggsave(filename = "/SSFA_Lithics_plot.pdf", path = dir_out, 
       device = "pdf", width = 190, height = 240, units = "mm")
```


## Zoom in for Smfc
The previous plots for the parameter Smfc show some extreme values for the sheep and lithic datasets.  
In order to better visualize the differences in software, we need zoomed-in plots excluding these extreme values (> 10).

```{r}
y_Smfc <- "Smfc"
ext_val <- 10

# Guinea pig dataset
GP_plot_Smfc <- filter(GP_plot, name == y_Smfc)
p_GP_Smfc <- ggplot(GP_plot_Smfc, 
                    aes_string(x = x_var_GP, y = "value", color = grp_colors)) +
             geom_boxplot(outlier.shape = NA) + 
             geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
             labs(x = NULL, y = y_Smfc) + 
             theme_classic() +
             theme(legend.position="bottom", legend.title = element_blank())
print(p_GP_Smfc)
ggsave(plot = p_GP_Smfc, filename = "/SSFA_GuineaPigs_plot_Smfc.pdf", path = dir_out, 
       device = "pdf")

# Sheep dataset
sheep_plot_Smfc <- filter(sheep_plot, name == y_Smfc)
sheep_plot_Smfc_filt <- filter(sheep_plot_Smfc, value <= ext_val)
p_sheep_Smfc <- ggplot(sheep_plot_Smfc_filt, 
                       aes_string(x = x_var_sheep, y = "value", color = grp_colors)) +
                geom_boxplot(outlier.shape = NA) + 
                geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
                labs(x = NULL, y = y_Smfc) + 
                theme_classic() +
                theme(legend.position="bottom", legend.title = element_blank())
print(p_sheep_Smfc)
ggsave(plot = p_sheep_Smfc, filename = "/SSFA_Sheeps_plot_Smfc.pdf", path = dir_out, 
       device = "pdf")

# Lithic dataset
lith_plot_Smfc <- filter(lith_plot, name == y_Smfc)
lith_plot_Smfc_filt <- filter(lith_plot_Smfc, value <= ext_val)
p_lith_Smfc <- ggplot(lith_plot_Smfc_filt, 
                      aes_string(x = x_var_lith, y = "value", color = grp_colors)) +
               geom_boxplot(outlier.shape = NA) + 
               geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
               facet_wrap(~Treatment, scale = "free_y") +
               labs(x = NULL, y = y_Smfc) + 
               theme_classic() +
               theme(legend.position="bottom", legend.title = element_blank())
print(p_lith_Smfc) 
ggsave(plot = p_lith_Smfc, filename = "/SSFA_Lithics_plot_Smfc.pdf", path = dir_out, 
       device = "pdf")
```

These plots of Smfc do not show all data points.  
For the sheep plot, these points are outside of the y-range shown and are therefore excluded from the plot:  
```{r}
data.frame(sheep_plot_Smfc[sheep_plot_Smfc$value > ext_val, ])
```

For the lithic plots, these points are outside of the y-ranges shown and were therefore excluded from the plots:   
```{r}
data.frame(lith_plot_Smfc[lith_plot_Smfc$value > ext_val, ])
```


---


# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("R_analysis/scripts/SSFA_0_RStudioVersion.txt", n = 1)`.


---


END OF SCRIPT
