---
title: "Import SSFA datasets"
author: "Ivan Calandra"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---


```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---


# Goal of the script
This script formats the output of the resulting files from applying SSFA.
The script will:

1. Read in the original files  
2. Format the data  
3. Write XLSX-files and save R objects ready for further analysis in R  

```{r}
dir_in  <- "R_analysis/raw_data"
dir_out <- "R_analysis/derived_data"
```

Raw data must be located in "`r paste0("~/", dir_in)`".  
Formatted data will be saved in "`r paste0("~/", dir_out)`".

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(openxlsx)
library(R.utils)
library(tools)
library(tidyverse)
```


---


# Get names, path and information of all files 
```{r}
info_in <- list.files(dir_in, pattern = "\\.csv|xlsx$", full.names = TRUE) %>% 
           md5sum()
```

The checksum (MD5 hashes) of the imported files are:  
```{r, echo = FALSE}
data.frame(files = basename(names(info_in)), checksum = info_in, row.names = NULL)
```


---


# Guinea pigs
## ConfoMap
### Read in data
```{r}
# Extract file name for ConfoMap analysis on Guinea Pigs, and read in file
confoGP <- names(info_in) %>% 
           .[grepl("ConfoMap", .) & grepl("GuineaPigs", .)] %>% 
           read.csv(header = FALSE, na.strings = "*****")
```

### Exclude repeated rows and select columns
```{r}
confoGP_keep_col  <- c(12, 46:52) # Define columns to keep
confoGP_keep_rows <- which(confoGP[[1]] != "#") # Define rows to keep
confoGP_keep      <- confoGP[confoGP_keep_rows, confoGP_keep_col] # Subset rows and columns
```

### Add headers
```{r}
# Get headers from 2nd row
colnames(confoGP_keep) <- confoGP[2, confoGP_keep_col] %>% 
  
                          # Convert to valid names
                          make.names() %>% 
  
                          # Delete repeated periods
                          gsub("\\.+", "\\.", x = .) %>% 
  
                          # Delete periods at the end of the names
                          gsub("\\.$", "", x = .) %>%
  
                          # Keep only part after the last period
                          gsub("^([A-Za-z0-9]+\\.)+", "", x = .) 
```

### Edit column "Name"
```{r}
# Delete everything from the beginning to " --- "
confoGP_keep[["Name"]] <- gsub("(^([_A-Za-z0-9-]+) --- )", "", confoGP_keep[["Name"]])
```

### Extract units
```{r}
confoGP_units <- unlist(confoGP[3, confoGP_keep_col[-1]]) # Extract unit line for considered columns
names(confoGP_units) <- colnames(confoGP_keep)[-1]        # Get names associated to the units
```

### Convert to numeric and add column about software
```{r}
confoGP_keep <- type_convert(confoGP_keep) %>% 
                mutate(Software = "ConfoMap")
```

### Edit column "Name"
```{r}
# Delete everything from the beginning to " --- "
confoGP_keep[["Name"]] <- gsub("(^([_A-Za-z0-9-]+) --- )", "", confoGP_keep[["Name"]])
```


## Toothfrax
### Read in data
```{r}
# Extract file name for Toothfrax analysis on Guinea Pigs
toothfraxGP <- names(info_in) %>% 
               .[grepl("Toothfrax", .) & grepl("GuineaPigs", .)] %>% 

               # Read in file
               read.xlsx(check.names = TRUE)
```

### Select columns, edit headers and add column about software
```{r}
toothfraxGP_keep <- select(toothfraxGP, Name = Filename, epLsar = epLsar1.80µm, 
                           R² = Rsquared, Asfc = Asfc1om, Smfc = LineStart, 
                           HAsfc9 = X3x3HAsfc, HAsfc81 = X9x9HAsfc) %>% 
                    mutate(Software = "Toothfrax")
```

### Edit column "Name"
```{r}
# Remove ".sur" at the end of the name and delete everything from the beginning to " --- "
toothfraxGP_keep[["Name"]] <- gsub("(\\.sur$)|(^([_A-Za-z0-9-]+) --- )", "", 
                                   toothfraxGP_keep[["Name"]])
```

### Get units
```{r}
# Check if all units are identical
if (length(unique(toothfraxGP[["Units"]])) != 1) {
  warning(paste("The studiables have different units for the SSFA parameters.",
                "Only the unit for the first studiable will be kept.",
                "Please check in the original file.", sep = "\n"))
}

# Get units from the column "Units"
toothfraxGP_units <- toothfraxGP[["Units"]][1]

# Add names to it
names(toothfraxGP_units) <- "Toothfrax_units" 
```


## Merge datasets
### Merge
```{r}
GP_keep <- merge(confoGP_keep, toothfraxGP_keep, all = TRUE)
```

### Add column about diet
```{r}
GP_keep[grep("2CC4", GP_keep[["Name"]]), "Diet"] <- "Dry lucerne"
GP_keep[grep("2CC5", GP_keep[["Name"]]), "Diet"] <- "Dry grass"
GP_keep[grep("2CC6", GP_keep[["Name"]]), "Diet"] <- "Dry bamboo"
```

### Add column about dataset and re-order columns
```{r}
GP_final <- GP_keep %>% 
            mutate(Dataset = "GuineaPigs") %>% 
            select(Dataset, Name, Diet, Software, everything())
```

### Combine both sets of units
```{r}
GP_units <- c(confoGP_units, toothfraxGP_units) 
```

### Check the result
```{r}
str(GP_final)
head(GP_final)
```


---


# Sheeps
Note that for compactness, comments to explain the code are given only in the section about [Guinea Pigs](#guinea-pigs).

## ConfoMap
### Read in data
```{r}
confoSheep <- names(info_in) %>% 
              .[grepl("ConfoMap", .) & grepl("Sheeps", .)] %>% 
              read.csv(header = FALSE, na.strings = "*****")
```

### Exclude repeated rows and select columns
```{r}
confoSheep_keep_col  <- c(12, 43:49)
confoSheep_keep_rows <- which(confoSheep[[1]] != "#")  
confoSheep_keep      <- confoSheep[confoSheep_keep_rows, confoSheep_keep_col]
```

### Add headers
```{r}
colnames(confoSheep_keep) <- confoSheep[2, confoSheep_keep_col] %>% 
                             make.names() %>% 
                             gsub("\\.+", "\\.", x = .) %>% 
                             gsub("\\.$", "", x = .) %>%
                             gsub("^([A-Za-z0-9]+\\.)+", "", x = .)
```

### Extract units
```{r}
confoSheep_units <- unlist(confoSheep[3, confoSheep_keep_col[-1]]) %>% 
                    gsub("Â", "", .) 
names(confoSheep_units) <- colnames(confoSheep_keep)[-1] 
```

### Convert to numeric and add column about software
```{r}
confoSheep_keep <- type_convert(confoSheep_keep) %>% 
                   mutate(Software = "ConfoMap")
```

### Edit column "Name"
```{r}
# Delete everything from the beginning to " --- "
confoSheep_keep[["Name"]] <- gsub("(^([_A-Za-z0-9-]+) --- )", "", confoSheep_keep[["Name"]])
```


## Toothfrax
### Read in data
```{r}
toothfraxSheep <- names(info_in) %>% 
                  .[grepl("Toothfrax", .) & grepl("Sheeps", .)] %>% 
                  read.xlsx(check.names = TRUE)
```

### Select columns, edit headers and add column about software
```{r}
toothfraxSheep_keep <- select(toothfraxSheep, Name = Filename, epLsar = epLsar1.80µm, 
                              R² = Rsquared, Asfc = Asfc1om, Smfc = LineStart, 
                              HAsfc9 = X3x3HAsfc, HAsfc81 = X9x9HAsfc) %>% 
                       mutate(Software = "Toothfrax")
```

### Edit column "Name"
```{r}
toothfraxSheep_keep[["Name"]] <- gsub("(\\.sur$)|(^([_A-Za-z0-9-]+) --- )", "", 
                                      toothfraxSheep_keep[["Name"]]) 
```

### Get units
```{r}
if (length(unique(toothfraxSheep[["Units"]])) != 1) {
  warning(paste("The studiables have different units for the SSFA parameters.",
                "Only the unit for the first studiable will be kept.",
                "Please check in the original file.", sep = "\n"))
} 
toothfraxSheep_units <- toothfraxSheep[["Units"]][1] 
names(toothfraxSheep_units) <- "Toothfrax_units" 
```


## Merge datasets
### Merge
```{r}
sheep_keep <- merge(confoSheep_keep, toothfraxSheep_keep, all = TRUE)
```

### Add column about diet
```{r}
sheep_keep[grep("L5", sheep_keep[["Name"]]), "Diet"] <- "Clover"
sheep_keep[grep("L6", sheep_keep[["Name"]]), "Diet"] <- "Clover+dust"
sheep_keep[grep("L7", sheep_keep[["Name"]]), "Diet"] <- "Grass"
sheep_keep[grep("L8", sheep_keep[["Name"]]), "Diet"] <- "Grass+dust"
```

### Add column about dataset and re-order columns
```{r}
sheep_final <- sheep_keep %>% 
               mutate(Dataset = "Sheeps") %>% 
               select(Dataset, Name, Diet, Software, everything())
```

### Add units
```{r}
sheep_units <- c(confoSheep_units, toothfraxSheep_units) 
```

### Check the result
```{r}
str(sheep_final)
head(sheep_final)
```


---


# Lithics
Note that for compactness, comments to explain the code are given only in the section about [Guinea Pigs](#guinea-pigs).

## ConfoMap
### Read in data
```{r}
confoLith <- names(info_in) %>% 
             .[grepl("ConfoMap", .) & grepl("Lithics", .)] %>% 
             read.csv(header = FALSE, na.strings = "*****")
```

### Exclude repeated rows and select columns
```{r}
confoLith_keep_col  <- c(12, 47:53) 
confoLith_keep_rows <- which(confoLith[[1]] != "#")
confoLith_keep      <- confoLith[confoLith_keep_rows, confoLith_keep_col] 
```

### Add headers
```{r}
colnames(confoLith_keep) <- confoLith[2, confoLith_keep_col] %>% 
                            make.names() %>% 
                            gsub("\\.+", "\\.", x = .) %>% 
                            gsub("\\.$", "", x = .) %>%
                            gsub("^([A-Za-z0-9]+\\.)+", "", x = .)
```

### Extract units
```{r}
confoLith_units <- unlist(confoLith[3, confoLith_keep_col[-1]]) 
names(confoLith_units) <- colnames(confoLith_keep)[-1]
```

### Convert to numeric and add column about software
```{r}
confoLith_keep <- type_convert(confoLith_keep) %>% 
                  mutate(Software = "ConfoMap")
```

### Edit column "Name"
```{r}
# Delete everything from the beginning to " --- "
confoLith_keep[["Name"]] <- gsub("(^([_A-Za-z0-9-]+) --- )", "", confoLith_keep[["Name"]])
```


## Toothfrax
### Read in data
```{r}
toothfraxLith <- names(info_in) %>% 
                 .[grepl("Toothfrax", .) & grepl("Lithics", .)] %>% 
                 read.xlsx(check.names = TRUE)
```

### Select columns, edit headers and add column about software
```{r}
toothfraxLith_keep <- select(toothfraxLith, Name = Filename, epLsar = epLsar1.80µm, 
                             R² = Rsquared, Asfc = Asfc1om, Smfc = LineStart, 
                             HAsfc9 = X3x3HAsfc, HAsfc81 = X9x9HAsfc) %>% 
                      mutate(Software = "Toothfrax")
```

### Edit column "Name"
```{r}
toothfraxLith_keep[["Name"]] <- gsub("(\\.sur$)|(^([_A-Za-z0-9-]+) --- )", "", 
                                     toothfraxLith_keep[["Name"]])
```

### Get units
```{r}
if (length(unique(toothfraxLith[["Units"]])) != 1) {
  warning(paste("The studiables have different units for the SSFA parameters.",
                "Only the unit for the first studiable will be kept.",
                "Please check in the original file.", sep = "\n"))
}  
toothfraxLith_units <- toothfraxLith[["Units"]][1]
names(toothfraxLith_units) <- "Toothfrax_units"
```


## Merge datasets
### Merge
```{r}
lith_keep <- merge(confoLith_keep, toothfraxLith_keep, all = TRUE)
```

### Exclude FLT3-8_Area1 ("before" wrongly acquired )
```{r}
FLT3_8_Area1 <- grep("^FLT3-8_.+_Area1", lith_keep[["Name"]])
lith_keep    <- lith_keep[-FLT3_8_Area1, ]
```

### Add columns about before/after and treatment
```{r}
lith_keep[grep("LSM_",  lith_keep[["Name"]]), "Before.after"] <- "Before"
lith_keep[grep("LSM2_", lith_keep[["Name"]]), "Before.after"] <- "After"
lith_keep[["Before.after"]] <- factor(lith_keep[["Before.after"]], levels = c("Before", "After"))

lith_keep[grep("FLT3-8|QTZ3-2",   lith_keep[["Name"]]), "Treatment"] <- "Control"
lith_keep[grep("FLT3-9|QTZ3-5",   lith_keep[["Name"]]), "Treatment"] <- "BrushNoDirt"
lith_keep[grep("FLT3-13|QTZ3-3",  lith_keep[["Name"]]), "Treatment"] <- "RubDirt"
lith_keep[grep("FLT3-10|QTZ3-13", lith_keep[["Name"]]), "Treatment"] <- "BrushDirt"
lith_keep[["Treatment"]] <- factor(lith_keep[["Treatment"]], 
                                   levels = c("Control", "RubDirt", "BrushNoDirt", "BrushDirt"))
```

### Add column about dataset and re-order columns
```{r}
lith_final <- lith_keep %>% 
              mutate(Dataset = "Lithics") %>% 
              select(Dataset, Name, Treatment, Before.after, Software, everything())
```

### Add units
```{r}
lith_units <- c(confoLith_units, toothfraxLith_units) 
```

### Check the result
```{r}
str(lith_final)
head(lith_final)
```


---


# Combine datasets
## Merge datasets
```{r}
# Merge the first two datasets
all_data <- merge(GP_final, sheep_final, all = TRUE) %>% 
  
            # Merge the first two with the third dataset
            merge(lith_final, all = TRUE) %>% 
  
            # Re-order columns
            select(Dataset, Name, Software, Diet, Treatment, Before.after, everything()) %>% 
  
            # Re-order rows
            arrange(Dataset, Name, Software, Diet, Treatment, Before.after)
```


## Combine units
```{r}
# Test if units are identical across datasets
if (all(sapply(list(GP_units, sheep_units), FUN = identical, lith_units)) == TRUE) {
  
  # If TRUE, use only lith_units
  # Combine into a data.frame for export
  units_table <- data.frame(variable = names(lith_units), units = lith_units) 
  # Add as comment
  comment(all_data) <- lith_units
  
  # If FALSE, use all units
} else {
  
  # Combine into a data.frame for export
  units_table <- data.frame(variable = names(lith_units), units_GP = GP_units, 
                            units_sheep = sheep_units, units_lith = lith_units) 
  
  # Combine all units into a single vector
  units_all <- c(GP_units, sheep_units, lith_units)
  
  # Rename the vector to reflect the dataset
  names(units_all) <- length(GP_units) %>% 
                      rep(c("GP", "sheep", "lith"), each = .) %>% 
                      paste(names(x), sep=".")
  
  # Add as comment
  comment(all_data) <- units_all
}
```

Type `comment(all_data)` to check the units of the parameters.


## Check the result
```{r}
str(all_data)
head(all_data)
```


---


# Save data
## Create file names
```{r}
all_xlsx <- paste0(dir_out, "/SSFA_all_data.xlsx")
all_rbin <- paste0(dir_out, "/SSFA_all_data.Rbin")
```


## Write to XLSX and Rbin
```{r}
write.xlsx(list(data = all_data, units = units_table), file = all_xlsx) 
saveObject(all_data, file = all_rbin) 
```

Rbin files (e.g. `SSFA_all_data.Rbin`) can be easily read into an R object (e.g. `rbin_data`) using the following code:
```{r, eval = FALSE}
library(R.utils)
rbin_data <- loadObject("SSFA_all_data.Rbin")
```


---


# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("R_analysis/scripts/SSFA_0_RStudioVersion.txt", n = 1)`.


---


END OF SCRIPT
