---
title: "Plots of SSFA variables"
author: "Ivan Calandra"
date: "`r Sys.time()`"
output:
  github_document: 
    toc: true
    toc_depth: 2
    html_preview: false
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---


# Goal of the script
The script plots all SSFA variables for each dataset to compare between the ConfoMap and Toothfrax analyses.   

```{r}
dir_in  <- "R_analysis/derived_data"
dir_out <- "R_analysis/plots"
```

Input Rbin data file must be located in "`r paste0("~/", dir_in)`".  
Plots will be saved in "`r paste0("~/", dir_out)`".

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
```


---


# Read in data
## Get names, path and information of input file 
```{r}
info_in <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE) %>% 
           md5sum()
```

The checksum (MD5 hashes) of the loaded file is:  
```{r, echo = FALSE}
data.frame(files = basename(names(info_in)), checksum = info_in, row.names = NULL)
```


## Read in Rbin file
```{r}
all_data <- loadObject(names(info_in))
str(all_data)
```


---


# Define variables
```{r}
# x-axis (grouping)
x_var_GP <- x_var_sheep <- "Diet"
x_var_lith <- "Before.after"

# y-axis
y_var <- c("Asfc", "Smfc", "epLsar", "NewEplsar", "HAsfc9", "HAsfc81", "R²")

# colors
grp_colors <- "Software"

# subplots Lithics dataset
facet_lith <- "Treatment"
```

The following variables will be used:  
```{r}
x_var_GP
x_var_sheep
x_var_lith
grp_colors
facet_lith
y_var
```


---


# Plot each set of the selected numeric variables
## Guinea Pigs
```{r}
# Filter dataset
GP <- filter(all_data, Dataset == "GuineaPigs")

# Change from wide to long format
GP_plot <- pivot_longer(GP[c(x_var_GP, grp_colors, y_var)], all_of(y_var))

# Re-order factor levels to fit order of plots on facet
GP_plot$name <- factor(GP_plot$name, levels = y_var)

# Plot all variables at once using facet_wrap()
p_GP <- ggplot(GP_plot, aes_string(x = x_var_GP, y = "value", color = grp_colors)) +
        geom_boxplot(outlier.shape = NA) + 
        geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
        facet_wrap(~name, scales = "free_y", ncol = 2) +
        labs(x = NULL, y = NULL) + 
        theme_classic()

# Position the legend into an empty facet
p_GP <- p_GP + theme(legend.position = c(1, 0), legend.justification = c(1, 0))

# Print and save resulting plot
print(p_GP)
ggsave(filename = "/SSFA_GuineaPigs_plot.pdf", path = dir_out, 
       device = "pdf", width = 210, height = 297, units = "mm")
```

Warnings are due to the missing values for Toothfrax - NewEplsar.


## Sheeps
Note that for compactness, comments to explain the code are given only in the section about [Guinea Pigs](#guinea-pigs).

```{r}
sheep <- filter(all_data, Dataset == "Sheeps")
sheep_plot <- pivot_longer(sheep[c(x_var_sheep, grp_colors, y_var)], all_of(y_var))
sheep_plot$name <- factor(sheep_plot$name, levels = y_var)
p_sheep <- ggplot(sheep_plot, aes_string(x = x_var_sheep, y = "value", color = grp_colors)) +
           geom_boxplot(outlier.shape = NA) + 
           geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
           facet_wrap(~name, scales = "free_y", ncol = 2) +
           labs(x = NULL, y = NULL) + 
           theme_classic() +
           theme(legend.position = c(1, 0), legend.justification = c(1, 0))
print(p_sheep)
ggsave(filename = "/SSFA_Sheeps_plot.pdf", path = dir_out, 
       device = "pdf", width = 210, height = 297, units = "mm")
```


## Lithics
Note that for compactness, comments to explain the code are given only in the section about [Guinea Pigs](#guinea-pigs). 

There is one difference though: here, three columns are used for the grouping ("Software", "Treatment", and "Before.after").  
Software is still shown with colors.  
"Before.after" is plotted on the x-axis and the variables are on the y-axes.  
`facet_grid()` is used to plot a grid of subplots, with variables in rows and treatments in columns.

```{r}
lith <- filter(all_data, Dataset == "Lithics")
lith_plot <- pivot_longer(lith[c(x_var_lith, grp_colors, facet_lith, y_var)], all_of(y_var))
lith_plot$name <- factor(lith_plot$name, levels = y_var)
p_lith <- ggplot(lith_plot, aes_string(x = x_var_lith, y = "value", color = grp_colors)) +
          geom_boxplot(outlier.shape = NA) + 
          geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
          facet_grid(as.formula(paste0("name~", facet_lith)), scales = "free_y") +
          labs(x = NULL, y = NULL) + 
          theme_classic()
print(p_lith)
ggsave(filename = "/SSFA_Lithics_plot.pdf", path = dir_out, 
       device = "pdf", width = 210, height = 297, units = "mm")
```


---


# Show plot files information
```{r}
info_out <- list.files(path = dir_out, pattern = "\\.pdf$", full.names = TRUE) %>% 
            md5sum()
```

The following plot(s) was (were) created and its (their) checksum(s) is (are):
```{r, echo = FALSE}
data.frame(files = basename(names(info_out)), checksum = info_out, row.names = NULL)
```


---


# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("R_analysis/scripts/RStudioVersion.txt", n = 1)`.


---


END OF SCRIPT
