---
title: "Plots of SSFA variables"
author: "Ivan Calandra"
date: "`r Sys.time()`"
output:
  github_document: 
    toc: true
    toc_depth: 2
    html_preview: false
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---


# Goal of the script
The script plots all SSFA variables for each dataset to compare between the ConfoMap and Toothfrax analyses.   

```{r}
dir_in  <- "R_analysis/derived_data"
dir_out <- "R_analysis/plots"
```

Input Rbin data file must be located in "`r paste0("~/", dir_in)`".  
Plots will be saved in "`r paste0("~/", dir_out)`".

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(ggh4x)
```


---


# Read in data
## Get names, path and information of input file 
```{r}
info_in <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE) %>% 
           md5sum()
```

The checksum (MD5 hashes) of the loaded file is:  
```{r, echo = FALSE}
data.frame(files = basename(names(info_in)), checksum = info_in, row.names = NULL)
```


## Read in Rbin file
```{r}
all_data <- loadObject(names(info_in))
str(all_data)
```


---


# Define variables
```{r}
# x-axis (grouping)
x_var_GP <- x_var_sheep <- "Diet"
x_var_lith <- "Before.after"

# y-axis
y_var <- c("Asfc", "Smfc", "HAsfc9", "HAsfc81", "epLsar", "NewEplsar")

# colors
grp_colors <- "Software"

# subplots Lithics dataset
facet_lith <- "Treatment"
```

The following variables will be used:  
```{r}
x_var_GP
x_var_sheep
x_var_lith
grp_colors
facet_lith
y_var
```


---


# Calculate y-scales
The range of the y-scales on the plots should be the same for the guinea pig and sheep datasets. Lithics are not comparable at all so this dataset is plotted with appropriate y-scales for this dataset alone.

```{r}
# Select guinea pig and sheep datasets
GP_sheep <- filter(all_data, Dataset %in% c("GuineaPigs", "Sheeps"))

# Create a named empty list to store the ranges of each parameter
yscales <- vector(mode = "list", length = length(y_var))
names(yscales) <- y_var

# Calculate the range of each parameter
for (i in y_var) {
  yscales[[i]] <- scale_y_continuous(limits = range(GP_sheep[[i]], na.rm = TRUE))
}
```

---


# Plot each set of the selected numeric variables
## Guinea Pigs
```{r}
# Filter dataset
GP <- filter(all_data, Dataset == "GuineaPigs")

# Change from wide to long format
GP_plot <- pivot_longer(GP[c(x_var_GP, grp_colors, y_var)], all_of(y_var))

# Re-order factor levels to fit order of plots on facet
GP_plot$name <- factor(GP_plot$name, levels = y_var)

# Plot all variables at once using facet_wrap()
p_GP <- ggplot(GP_plot, aes_string(x = x_var_GP, y = "value", color = grp_colors)) +
        
        # Do not show outliers
        geom_boxplot(outlier.shape = NA) + 
  
        # Define jitter for points within boxplots
        geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
  
        # Wrap around parameters
        facet_wrap(~name, scales = "free_y", ncol = 2) +
  
        # Use custom y-scales
        facetted_pos_scales(y = yscales) +
  
        # Remove axis labels
        labs(x = NULL, y = NULL) + 

        # Add an empty level to the x-axis to have the same with as for the sheep dataset
        scale_x_discrete(limits = c(unique(GP_plot$Diet), " ")) +

        # Choose a light theme
        theme_classic()

# Print and save resulting plot
print(p_GP)
ggsave(filename = "/SSFA_GuineaPigs_plot.pdf", path = dir_out, 
       device = "pdf", width = 210, height = 297, units = "mm")
```

Warnings are due to the missing values for Toothfrax - NewEplsar.


## Sheeps
Note that for compactness, comments to explain the code are given only in the section about [Guinea Pigs](#guinea-pigs).

```{r}
sheep <- filter(all_data, Dataset == "Sheeps")
sheep_plot <- pivot_longer(sheep[c(x_var_sheep, grp_colors, y_var)], all_of(y_var))
sheep_plot$name <- factor(sheep_plot$name, levels = y_var)
p_sheep <- ggplot(sheep_plot, aes_string(x = x_var_sheep, y = "value", color = grp_colors)) +
           geom_boxplot(outlier.shape = NA) + 
           geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
           facet_wrap(~name, scales = "free_y", ncol = 2) +
           facetted_pos_scales(y = yscales) +
           labs(x = NULL, y = NULL) + 
           theme_classic()
print(p_sheep)
ggsave(filename = "/SSFA_Sheeps_plot.pdf", path = dir_out, 
       device = "pdf", width = 210, height = 297, units = "mm")
```


## Lithics
Note that for compactness, comments to explain the code are given only in the section about [Guinea Pigs](#guinea-pigs). 

There is one difference though: here, three columns are used for the grouping ("Software", "Treatment", and "Before.after").  
Software is still shown with colors.  
"Before.after" is plotted on the x-axis and the variables are on the y-axes.  
`facet_grid()` is used to plot a grid of subplots, with variables in rows and treatments in columns.

```{r}
lith <- filter(all_data, Dataset == "Lithics")
lith_plot <- pivot_longer(lith[c(x_var_lith, grp_colors, facet_lith, y_var)], all_of(y_var))
lith_plot$name <- factor(lith_plot$name, levels = y_var)
p_lith <- ggplot(lith_plot, aes_string(x = x_var_lith, y = "value", color = grp_colors)) +
          geom_boxplot(outlier.shape = NA) + 
          geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
          facet_grid(as.formula(paste0("name~", facet_lith)), scales = "free_y") +
          labs(x = NULL, y = NULL) + 
          theme_classic()
print(p_lith)
ggsave(filename = "/SSFA_Lithics_plot.pdf", path = dir_out, 
       device = "pdf", width = 210, height = 297, units = "mm")
```


## Zoom in for Smfc
The previous plots for the parameter Smfc show some extreme values.  
In order to better visualize the differences in software, we need zoomed-in plots excluding these extreme values.

```{r}
y_Smfc <- "Smfc"

# Guinea pig dataset
GP_plot_Smfc <- filter(GP_plot, name == y_Smfc)
p_GP_Smfc <- ggplot(GP_plot_Smfc, 
                    aes_string(x = x_var_GP, y = "value", color = grp_colors)) +
             geom_boxplot(outlier.shape = NA) + 
             geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
             labs(x = NULL, y = y_Smfc) + 
             theme_classic()
print(p_GP_Smfc)
ggsave(plot = p_GP_Smfc, filename = "/SSFA_GuineaPigs_plot_Smfc.pdf", path = dir_out, 
       device = "pdf")

# Sheep dataset
sheep_plot_Smfc <- filter(sheep_plot, name == y_Smfc)
sheep_plot_Smfc_10 <- filter(sheep_plot_Smfc, value <= 10)
p_sheep_Smfc <- ggplot(sheep_plot_Smfc_10, 
                       aes_string(x = x_var_sheep, y = "value", color = grp_colors)) +
                geom_boxplot(outlier.shape = NA) + 
                geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
                labs(x = NULL, y = y_Smfc) + 
                theme_classic()
print(p_sheep_Smfc)
ggsave(plot = p_sheep_Smfc, filename = "/SSFA_Sheeps_plot_Smfc.pdf", path = dir_out, 
       device = "pdf")

# Lithic dataset
lith_plot_Smfc <- filter(lith_plot, name == y_Smfc)
lith_plot_Smfc_5 <- filter(lith_plot_Smfc, value <= 5)
p_lith_Smfc <- ggplot(lith_plot_Smfc_5, 
                      aes_string(x = x_var_lith, y = "value", color = grp_colors)) +
               geom_boxplot(outlier.shape = NA) + 
               geom_point(position = position_jitterdodge(jitter.width = 0.2, seed = 123)) +
               facet_wrap(~Treatment, scale = "free_y") +
               labs(x = NULL, y = y_Smfc) + 
               theme_classic()
print(p_lith_Smfc) 
ggsave(plot = p_lith_Smfc, filename = "/SSFA_Lithics_plot_Smfc.pdf", path = dir_out, 
       device = "pdf")
```

These plots of Smfc do not show all data points.  
For the sheep plot, `r length(which(sheep_plot_Smfc$value > 10))` points are outside of the y-range shown and are therefore excluded from the plot.  
For the lithic plots, `r length(which(lith_plot_Smfc$value > 5))` points are outside of the y-ranges shown and are therefore excluded from the plots.   


---


# Show plot files information
```{r}
info_out <- list.files(path = dir_out, pattern = "\\.pdf$", full.names = TRUE) %>% 
            md5sum()
```

The following plot(s) was (were) created and its (their) checksum(s) is (are):
```{r, echo = FALSE}
data.frame(files = basename(names(info_out)), checksum = info_out, row.names = NULL)
```


---


# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("R_analysis/scripts/RStudioVersion.txt", n = 1)`.


---


END OF SCRIPT
