---
title: "Summary statistics on SSFA datasets"
author: "Ivan Calandra"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 2
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---


```{r Knitr Options, include = FALSE}
	knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---


# Goal of the script
This script computes standard descriptive statistics for each group.  
The groups are based on:

* Diet (Guinea Pigs and Sheeps datasets)  
* Material and before/after experiment (Lithics dataset)

It computes the following statistics:  

* sample size (n = `length`)  
* smallest value (`min`)  
* largest value (`max`)
* mean  
* median  
* standard deviation (`sd`)

```{r}
dir_in  <- "R_analysis/derived_data"
dir_out <- "R_analysis/summary_stats"
```

Input Rbin data file must be located in "`r paste0("~/", dir_in)`".  
Summary statistics table will be saved in "`r paste0("~/", dir_out)`".

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(openxlsx)
library(R.utils)
library(tools)
library(doBy)
library(tidyverse)
```


---


# Read in data
## Get name and path of input file 
```{r}
info_in <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
info_in
```


## Read in Rbin file
```{r}
all_data <- loadObject(info_in)
str(all_data)
```


---


# Summary statistics
## Create function to compute the statistics at once
```{r}
nminmaxmeanmedsd <- function(x){
	y <- x[!is.na(x)]     # Exclude NAs
	n_test <- length(y)   # Sample size (n)
	min_test <- min(y)    # Minimum
	max_test <- max(y)    # Maximum
	mean_test <- mean(y)  # Mean
 	med_test <- median(y) # Median
 	sd_test <- sd(y)      # Standard deviation
 	out <- c(n_test, min_test, max_test, mean_test, med_test, sd_test) # Concatenate
 	names(out) <- c("n", "min", "max", "mean", "median", "sd")         # Name values
 	return(out)                                                        # Object to return
}
```


## Define grouping and numerical variables to use
```{r}
# Define grouping variables
grp <- c("Dataset", "Software", "Diet", "Treatment", "Before.after")
```

The following grouping variables will be used: 
```{r, echo = FALSE}
cat(colnames(select(all_data, all_of(grp))), sep="\n")
```

All numerical variables will be used: 
```{r, echo = FALSE}
cat(colnames(select(all_data, where(is.numeric))), sep="\n")
```


## Subset out surfaces with too many non-measured points (NMP)
```{r}
data_nmp10   <- filter(all_data, NMP >= 10)
data_nmp5_10 <- filter(all_data, NMP < 10 & NMP >= 5)
data_nmp0_5  <- filter(all_data, NMP < 5)
```


## Compute summary statistics
```{r}
# All data
stats_all <- as.formula(paste(".~", paste(grp, collapse = "+"))) %>% 
             summaryBy(data = all_data, FUN = nminmaxmeanmedsd)

# NMP < 10%
stats_nmp10 <- as.formula(paste(".~", paste(grp, collapse = "+"))) %>% 
               summaryBy(data = data_nmp10, FUN = nminmaxmeanmedsd)

# NMP < 5%
stats_nmp5 <- as.formula(paste(".~", paste(grp, collapse = "+"))) %>% 
              summaryBy(data = data_nmp5, FUN = nminmaxmeanmedsd)
```

The warnings are due to the missing values for NewEplsar - Toothfrax. NewEplsar can only be computed in ConfoMap.
```{r}
str(stats_all)
str(stats_nmp10)
str(stats_nmp5)
```


---


# Write results to XLSX
```{r}
write.xlsx(list(all_data = stats_all, NMP10 = stats_nmp10, NMP5 = stats_nmp5), 
           file = paste0(dir_out, "/SSFA_summary_stats.xlsx"))
```


---


# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("R_analysis/scripts/SSFA_0_RStudioVersion.txt", n = 1)`.


---


END OF SCRIPT
